Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-07-27 paper/manuscript.tex
last changes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%
\begin{frontmatter}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Alchemical free energy calculations are a useful tool for predicting free energy differences associated with the transfer of molecules from one environment to another.
The hallmark of these methods is the use of "bridging" potential energy functions representing \emph{alchemical} intermediate states that cannot exist in as real chemical species. The data collected from these bridging alchemical thermodynamic states allows the efficient computation of transfer free energies (or differences in transfer free energies) with orders of magnitude less simulation time than simulating the transfer process directly.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Benchmark information  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alchemical free energy datasets: an overview}
\label{sec:benchmark}
The following contains a non-exhaustive summary of alchemical free energy datasets that can serve as a starting point to review approaches or test new implementations. The field is moving towards a more standardised way of generating protein-ligand benchmark datasets and the progress of these efforts can be tracked here: \url{https://github.com/openforcefield/FE-Benchmarks-Best-Practices}. Currently lacking an exhaustive set of benchmark datasets, the review by Williams-Noonan et al.~\cite{williams-noonan2018free} contains an overview of recently published alchemical free energy studies. For comparison of FEP+ and Gromacs (using the AMBER99SB-ILDN and GAFF2 force field), cf. the recently published study by Pérez-Benito et al.~\cite{perez-benito2019predicting}.
An overview of further suggested benchmark sets can be found in the review by Mobley and Gilson~\cite{mobley2017predicting} or on \url{alchemistry.org}~\cite{alchemistry}. These include cyclodextrins, the Cytochrome C peroxidase (CCP) protein model binding site, thrombin and bromodomains as well as solvation benchmark sets~\cite{paliwal2011benchmark}. Please refere to ~\ref{tab:benchmarks}, for a small overview of datasets, what forcefields they used, and what the original study was it came from. 
\begin{table}
\caption{Selection of example datasets}
\begin{tabular}{cccc}
\textbf{Publication} & \textbf{Targets} & \textbf{Ligands} & \textbf{Force Field} \\
\hline
\multicolumn{4}{|c|}{D3R Grand Challenges~\cite{D3R}} \\
\hline
GC3~\cite{gaieb2019d3r} & 6 & 266 & various \\
GC2~\cite{gaieb2018d3r} & 1 & 102 & various \\
GC2015~\cite{gathiaka2016d3r} & 2 & 215 & various \\
\hline
\multicolumn{4}{|c|}{SAMPL Challenges~\cite{SAMPL}} \\
\hline
SAMPL6~\cite{rizzi2018overview} & 3 & 21 & various \\
SAMPL5~\cite{yin2017overview} & 3 & 22 & various \\
SAMPL4~\cite{muddana2014sampl4} & 2 & 23 & various \\
\hline
\multicolumn{4}{|c|}{Schrödinger Datasets} \\
\hline
FEP+ Dataset~\cite{wang2015accurate} & 8 & 199 & OPLS2.1 \\
FEP+ Dataset~\cite{harder2016opls3} & 8 & 199 & OPLS3 \\
FEP+ Dataset~\cite{roos2019opls3e} & 8 & 199 & OPLS3e \\
FEP+ Dataset~\cite{song2019using} & 8 & 199 & GAFF 1.8 \\
FEP+ Dataset~\cite{gapsys2020large} & 8 & 199 & various \\
FEP+ Dataset~\cite{kuhn2020assessment} & 8 & 199 & GAFF2.1 \\
Fragments ~\cite{steinbrecher2015accurate} & 8 & 96 & OPLS2.1 \\
Scaffold Hopping~\cite{wang2017accurate} & 6 & 21 & OPLS3 \\
Scaffold Hopping~\cite{kuhn2020assessment} & 6 & 21 & GAFF2.1 \\
Macrocycles~\cite{yu2017accurate} & 7 & 33 & OPLS3 \\
\hline
\multicolumn{4}{|c|}{Further Suggested Datasets} \\
\hline
Cucurbit[7]uril (CB7)~\cite{mobley2017predicting} & 1 & 15 & NA \\
Deep cavity cavitand~\cite{mobley2017predicting} & 2 & 19 & NA \\
T4 Lysozyme~\cite{mobley2017predicting} & 2 & 20 & NA \\
Merck set~\cite{MCompChem2019Sep} & 5 & 169 & OPSL3 \\

@@

\clearpage

@@

\section*{Author Contributions}
%%%%%%%%%%%%%%%%
% This section mustt describe the actual contributions of
% author. Since this is an electronic-only journal, there is
% no length limit when you describe the authors' contributions,
% so we recommend describing what they actually did rather than
% simply categorizing them in a small number of
% predefined roles as might be done in other journals.
%
% See the policies ``Policies on Authorship'' section of https://livecoms.github.io
% for more information on deciding on authorship and author order.
%%%%%%%%%%%%%%%%
\textbf{ASJSM}: Coordinated the document, contributed to most sections, and co-designed Figs.~\ref{fig:fig_binding_thermodynamic_cycle},~\ref{fig:fig_topology},~\ref{fig:fig_mcss},~\ref{fig:fig_types_of_networks},~\ref{fig:fig_absolute_thermodynamic_cycle},~\ref{fig:scatterplot_analysis},and created Figs.~\ref{fig:fig_what_is_lambda},~\ref{fig:fig_what_is_alchemy},~\ref{fig:overlap},~\ref[fig:pmf} and replotted~\ref{fig:automatic-equilibration-detection} and~\ref{fig:fig_types_of_networks}.\\
\textbf{JS}: Created Figs~ \ref{fig:fig_what_is_alchemy},~\ref{fig:fig_binding_thermodynamic_cycle},~\ref{fig:fig_topology},~\ref{fig:fig_mcss},~\ref{fig:fig_absolute_thermodynamic_cycle},~\ref{fig:scatterplot_analysis}, and an initial draft of~\ref{fig:fig_types_of_networks}. Wrote Sec.~\ref{sec:plot_data}, the checklist Sec.~\ref{sec:checklist}, and contributed to general formatting discussions and editing.\\
\textbf{MK}: Contributed to Sec.~\ref{sec:data_analysis}, provided the data for figure~\ref{fig:overlap}, compiled the dataset for Sec.~\ref{sec:benchmark} and helped edit the paper.\\
\textbf{DLM}: Contributed to the outline, drafted some of the sections, gave ideas on figures, and helped edit the paper.\\
\textbf{GT}: Contributed to Sec.~\ref{sec:intro} and~\ref{sec:drugdiscovery}, and helped edit the paper.\\
\textbf{AR}: Created figure~\ref{fig:freeenergytrajectories}, contributed to sections~\ref{sec:theory} and~\ref{sec:simulation_protocol_choice}, and helped edit the paper.\\
\textbf{MRS}: Helped create figure~\ref{fig:fig_what_is_lambda}, wrote Sec.~\ref{sec:important_path} describing choices for alchemical pathways and parts of~\ref{sec:data_analysis} on the analysis for free energy calculations. Reviewed and edited text throughout.\\
\textbf{LNN}: Helped write the simulation length, stopping conditions, and information saving section. Edited and reviewed alchemical path section.\\
\textbf{BA}: Helped write the uncertainty estimation, stopping conditions, and output analysis sections and created figure ~\ref{fig:convergence_forward_reverse}.\\
\textbf{JM}: Contributed to Sec.~\ref{subsec:reproducible},~\ref{sec:prerequisites},~\ref{sec:important_path},~\ref{subsec:estimators},~\ref{subsec:uncertainty}, and~\ref{sec:conclusion}\\
\textbf{JDC}: Wrote Sec.~\ref{sec:decorrelating-samples} and~\ref{sec:automatic-equilibration-detection} discussed structure and design of the whole document, suggested Figs.~\ref{fig:fig_what_is_alchemy} and~\ref{fig:fig_sampling_scheme} \\
\textbf{HX}: Contributed Sec.~\ref{subsec:accuracy}, to Sec.~\ref{sec:relative-fe-protocol}, and to Sec.~\ref{sec:are-they-good}.\\
\textbf{HBM} Contributed to Sec.~\ref{sec:plot_data} and Fig.~ \ref{fig:scatterplot_analysis} and helped edit the paper.\\
% We suggest you preserve this comment:
For a more detailed description of author contributions,
see the GitHub issue tracking and changelog at \githubrepository.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{frontmatter}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Alchemical free energy calculations are a useful tool for predicting free energy differences associated with the transfer of molecules from one environment to another.
The hallmark of these methods is the use of "bridging" potential energy functions representing \emph{alchemical} intermediate states that cannot exist in as real chemical species. The data collected from these bridging alchemical thermodynamic states allows the efficient computation of transfer free energies (or differences in transfer free energies) with orders of magnitude less simulation time than simulating the transfer process directly.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Benchmark information  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alchemical free energy datasets: an overview}
\label{sec:benchmark}
The following contains a non-exhaustive summary of alchemical free energy datasets that can serve as a starting point to review approaches or test new implementations. The field is moving towards a more standardised way of generating protein-ligand benchmark datasets and the progress of these efforts can be tracked here: \url{https://github.com/openforcefield/FE-Benchmarks-Best-Practices}. Currently lacking an exhaustive set of benchmark datasets, the review by Williams-Noonan et al.~\cite{williams-noonan2018free} contains an overview of recently published alchemical free energy studies. For comparison of FEP+ and Gromacs (using the AMBER99SB-ILDN and GAFF2 force field), cf. the recently published study by Pérez-Benito et al.~\cite{perez-benito2019predicting}.
An overview of further suggested benchmark sets can be found in the review by Mobley and Gilson~\cite{mobley2017predicting} or on \url{alchemistry.org}~\cite{alchemistry}. These include cyclodextrins, the Cytochrome C peroxidase (CCP) protein model binding site, thrombin and bromodomains as well as solvation benchmark sets~\cite{paliwal2011benchmark}. Please refere to ~\ref{tab:benchmarks}, for a small overview of datasets, what forcefields they used, and what the original study was it came from.

\begin{table}
\caption{Selection of example datasets}
\begin{tabular}{cccc}
\textbf{Publication} & \textbf{Targets} & \textbf{Ligands} & \textbf{Force Field} \\
\hline
\multicolumn{4}{|c|}{D3R Grand Challenges~\cite{D3R}} \\
\hline
GC3~\cite{gaieb2019d3r} & 6 & 266 & various \\
GC2~\cite{gaieb2018d3r} & 1 & 102 & various \\
GC2015~\cite{gathiaka2016d3r} & 2 & 215 & various \\
\hline
\multicolumn{4}{|c|}{SAMPL Challenges~\cite{SAMPL}} \\
\hline
SAMPL6~\cite{rizzi2018overview} & 3 & 21 & various \\
SAMPL5~\cite{yin2017overview} & 3 & 22 & various \\
SAMPL4~\cite{muddana2014sampl4} & 2 & 23 & various \\
\hline
\multicolumn{4}{|c|}{Schrödinger Datasets} \\
\hline
FEP+ Dataset~\cite{wang2015accurate} & 8 & 199 & OPLS2.1 \\
FEP+ Dataset~\cite{harder2016opls3} & 8 & 199 & OPLS3 \\
FEP+ Dataset~\cite{roos2019opls3e} & 8 & 199 & OPLS3e \\
FEP+ Dataset~\cite{song2019using} & 8 & 199 & GAFF 1.8 \\
FEP+ Dataset~\cite{gapsys2020large} & 8 & 199 & various \\
FEP+ Dataset~\cite{kuhn2020assessment} & 8 & 199 & GAFF2.1 \\
Fragments ~\cite{steinbrecher2015accurate} & 8 & 96 & OPLS2.1 \\
Scaffold Hopping~\cite{wang2017accurate} & 6 & 21 & OPLS3 \\
Scaffold Hopping~\cite{kuhn2020assessment} & 6 & 21 & GAFF2.1 \\
Macrocycles~\cite{yu2017accurate} & 7 & 33 & OPLS3 \\
\hline
\multicolumn{4}{|c|}{Further Suggested Datasets} \\
\hline
Cucurbit[7]uril (CB7)~\cite{mobley2017predicting} & 1 & 15 & NA \\
Deep cavity cavitand~\cite{mobley2017predicting} & 2 & 19 & NA \\
T4 Lysozyme~\cite{mobley2017predicting} & 2 & 20 & NA \\
Merck set~\cite{MCompChem2019Sep} & 5 & 169 & OPSL3 \\

@@

@@

\section*{Author Contributions}
%%%%%%%%%%%%%%%%
% This section mustt describe the actual contributions of
% author. Since this is an electronic-only journal, there is
% no length limit when you describe the authors' contributions,
% so we recommend describing what they actually did rather than
% simply categorizing them in a small number of
% predefined roles as might be done in other journals.
%
% See the policies ``Policies on Authorship'' section of https://livecoms.github.io
% for more information on deciding on authorship and author order.
%%%%%%%%%%%%%%%%
\textbf{ASJSM}: Coordinated the document, contributed to most sections, and co-designed Figs.~\ref{fig:fig_binding_thermodynamic_cycle},~\ref{fig:fig_topology},~\ref{fig:fig_mcss},~\ref{fig:fig_types_of_networks},~\ref{fig:fig_absolute_thermodynamic_cycle},~\ref{fig:scatterplot_analysis},and created Figs.~\ref{fig:fig_what_is_lambda},~\ref{fig:fig_what_is_alchemy},~\ref{fig:overlap},~\ref{fig:pmf} and replotted~\ref{fig:automatic-equilibration-detection} and~\ref{fig:fig_types_of_networks}.\\
\textbf{JS}: Created Figs~ \ref{fig:fig_what_is_alchemy},~\ref{fig:fig_binding_thermodynamic_cycle},~\ref{fig:fig_topology},~\ref{fig:fig_mcss},~\ref{fig:fig_absolute_thermodynamic_cycle},~\ref{fig:scatterplot_analysis}, and an initial draft of~\ref{fig:fig_types_of_networks}. Wrote Sec.~\ref{sec:plot_data}, the checklist Sec.~\ref{sec:checklist}, and contributed to general formatting discussions and editing.\\
\textbf{MK}: Contributed to Sec.~\ref{sec:data_analysis}, provided the data for figure~\ref{fig:overlap}, compiled the dataset for Sec.~\ref{sec:benchmark} and helped edit the paper.\\
\textbf{DLM}: Contributed to the outline, drafted some of the sections, gave ideas on figures, and helped edit the paper.\\
\textbf{GT}: Contributed to Sec.~\ref{sec:intro} and~\ref{sec:drugdiscovery}, and helped edit the paper.\\
\textbf{AR}: Created figure~\ref{fig:freeenergytrajectories}, contributed to sections~\ref{sec:theory} and~\ref{sec:simulation_protocol_choice}, and helped edit the paper.\\
\textbf{MRS}: Helped create figure~\ref{fig:fig_what_is_lambda}, wrote Sec.~\ref{sec:important_path} describing choices for alchemical pathways and parts of~\ref{sec:data_analysis} on the analysis for free energy calculations. Reviewed and edited text throughout.\\
\textbf{LNN}: Helped write the simulation length, stopping conditions, and information saving section. Edited and reviewed alchemical path section.\\
\textbf{BA}: Helped write the uncertainty estimation, stopping conditions, and output analysis sections and created figure ~\ref{fig:convergence_forward_reverse}.\\
\textbf{JM}: Contributed to Sec.~\ref{subsec:reproducible},~\ref{sec:prerequisites},~\ref{sec:important_path},~\ref{subsec:estimators},~\ref{subsec:uncertainty}, and~\ref{sec:conclusion}\\
\textbf{JDC}: Wrote Sec.~\ref{sec:decorrelating-samples} and~\ref{sec:automatic-equilibration-detection} discussed structure and design of the whole document, suggested Figs.~\ref{fig:fig_what_is_alchemy} and~\ref{fig:fig_sampling_scheme} \\
\textbf{HX}: Contributed Sec.~\ref{subsec:accuracy}, to Sec.~\ref{sec:relative-fe-protocol}, and to Sec.~\ref{sec:are-they-good}.\\
\textbf{HBM} Contributed to Sec.~\ref{sec:plot_data} and Fig.~ \ref{fig:scatterplot_analysis} and helped edit the paper.\\
% We suggest you preserve this comment:
For a more detailed description of author contributions,
see the GitHub issue tracking and changelog at \githubrepository.
